### 2024-12-03 manuscript.tex
Addressing comments

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). The high resolution allows for parameterizations of convection and
clouds to be avoided. Standard-resolution models have substantial cloud biases
over the Southern Ocean (SO), affecting radiation and sea surface temperature.
We evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Less stable conditions are the most problematic for
ICON and the reanalyses.  In daily cloud cover, ICON and the reanalyses tend to
be about 1 and 2 oktas clearer, respectively. Compared to radiosondes,
potential temperature is accurate in the reanalyses, but ICON underestimates
stability over the low-latitude SO and too humid in the boundary layer. MERRA-2
is too humid at all heights. SO cloud biases remain a substantial issue in the
GSRM, but are an improvement over the lower-resolution reanalyses. Explicitly
resolved convection and cloud processes were not enough to address the model
cloud biases.

@@

nextGEMS is a European Union--funded project \citep{nextgems} focused on the
research and development of GSRMs at multiple modelling centres and
universities in Europe.  The project also develops GSRM versions of the
Icosahedral Nonhydrostatic Weather and Climate Model (ICON), the Integrated
Forecasting System (IFS), and their ocean components at eddy-resolving
resolutions: ICON-O coupled with ICON and Finite-Element/volumE Sea ice-Ocean
Model (FESOM) and Nucleus for European Modelling of the Ocean (NEMO) coupled
with IFS.  The project has so far produced ICON and IFS simulations in three
cycles (Cycle 1--3) and pre-final simulation, with a final production
simulation planned by the end of the project. nextGEMS is not the only project
developing GSRMs. Other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM;
\cite{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated
with the cloud droplet size distribution, ice crystal habit, cloud lifetime,
and direct radiative interaction with aerosols.  In the 6$^\mathrm{th}$ phase
of the Coupled Model Intercomparison Project [CMIP6; \cite{eyring2016}], the
cloud feedback has increased relative to CMIP5 \citep{zelinka2020}, which is
one of the main reasons for the higher climate sensitivity of CMIP6 models.

@@

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but has too little cloudarea
overall. A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation by
approximately 10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5
km).  \cite{zhao2022} reported a similar SW radiation bias in five analysed
CMIP6 models over high-latitude SO and the total cloud fraction underestimation
on the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 (© Jeff Aquilina, Bureau of Meteorology).
}
\label{fig:map}
\end{figure}

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, the resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table
\ref{tab:voyages} list the voyages, campaigns, and stations, and Table
\ref{tab:voyage-references} lists references where available. Altogether, the
voyages and station dataset comprised 2421 days of data south of 40°S, but the
availability of ceilometer data was slightly smaller due to gaps in
measurements.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
operating at a wavelength of 910 nm, the Vaisala CT25K operating at 905 nm, and
the Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar principally designed to  measure cloud base, but they
also measure the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

The Vaisala CL51 (photo in Fig. \ref{fig:map}d) and CT25K are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
The CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K observations. The sampling (temporal) resolution is configurable, and in our
datasets is approximately 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied to the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results ch the observations better than the reanalyses in this metric.

@@

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including peak and total cloud
fraction. In these situations, the model vertical profiles of cloud occurrence compare well with observations, but still tend to
underestimate cloud occurrence above 1 km and near the surface. Non-cyclonic
situations are similar to the general case.

When subsetted by relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to relatively unstable situations, where the peak is more obtuse
and spread across the altitudes of 0--1 km.  In relatively stable
situations, the models do a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 underestimates cloud occurrence below 1 km (especially near the surface), and
MERRA-2 underestimates cloud occurrence even more strongly.

@@

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in Section \ref{sec:cloud-occurrence}. Observations display the highest proportion of
high cloud cover values (5--8 oktas), peaking at 7 oktas. This pattern is not represented by
ICON or either reanalyses.  While ICON is closest to matching the observed distribution, it tends to be 1 okta
clearer than the observations, peaking at 6 oktas, and highly underestimating
days with 8 oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

@@

We analysed about 2300 radiosonde profiles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and
temporally colocated profiles were taken from ICON and the reanalyses. Because
the time period of the ICON model output was different from the observations,
model time was chosen to be the same as the radiosonde launch time relative to
the start of the year. The profiles were partitioned into the same subsets as
above (Sections \ref{sec:cloud-occurrence} and \ref{sec:cloud-cover}).  Apart
from relative humidity, we focus on comparing virtual potential temperature
($\theta_v$) due to its role in low-level tropospheric stability, being one of
the primary factors affecting shallow convection and the associated low-level
cloud formation and dissipation. The observed and model profiles of virtual
potential temperature are shown in Fig. \ref{fig:potential-temperature}.

@@

The results imply that SO cloud biases are still a substantial issue even in the
km-scale resolution ICON model, even though an improvement over the
lower-resolution reanalyses is notable. More effort is needed to improve the
model cloud simulations in this fast-changing and understudied region. However, this analysis suggests that the
transition from models with parametrised convection and clouds to
storm-resolving models might not solve these biases without additional effort.
Evaluation of ocean--atmosphere heat, moisture, and momentum fluxes against in-situ observations over the SO and comparison of GSRM simulations against
large-eddy simulations are two potential avenues for future research that could
elucidate the physical mechanisms behind the biases, in addition to the more
common efforts in SO cloud microphysics and precipitation evaluation.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution allows for parameterizations of
convection and clouds to be avoided, relying instead on explicitly resolved
processes.  Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station during
2010--2021 using a ground-based lidar simulator.  ICON and the reanalyses
underestimate the total cloud fraction by about 10 and 20\%, respectively. ICON
and ERA5 overestimate the cloud occurrence peak at about 500 m, potentially
explained by their lifting condensation levels being too high.  The reanalyses
strongly underestimate near-surface clouds or fog.  MERRA-2 tends to
underestimate cloud occurrence at all heights. Less stable conditions are the
most problematic for ICON and the reanalyses.  In daily cloud cover, ICON and
the reanalyses tend to be about 1 and 2 oktas clearer, respectively. Compared
to radiosondes, potential temperature is accurate in the reanalyses, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes were not enough
to address the model cloud biases.

@@

nextGEMS is a project \citep{nextgems} focused on the
research and development of GSRMs at multiple modelling centres and
universities in Europe.  The project also develops GSRM versions of the
Icosahedral Nonhydrostatic Weather and Climate Model (ICON), the Integrated
Forecasting System (IFS), and their ocean components at eddy-resolving
resolutions: ICON-O coupled with ICON and Finite-Element/volumE Sea ice-Ocean
Model (FESOM) and Nucleus for European Modelling of the Ocean (NEMO) coupled
with IFS.  The project has so far produced ICON and IFS simulations with three
development versions called Cycle 1--3 and a pre-final version, with a final production
version planned by the end of the project. nextGEMS is not the only project
developing GSRMs; other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM;
\cite{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and both the indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated
with the cloud droplet size distribution, ice crystal habit, cloud lifetime,
and direct radiative interaction with aerosols.  In the 6$^\mathrm{th}$ phase
of the Coupled Model Intercomparison Project [CMIP6; \cite{eyring2016}], the
cloud feedback has increased relative to CMIP5 \citep{zelinka2020}, which is
one of the main reasons for the higher climate sensitivity of CMIP6 models.

@@

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models, which means that a model can maintain a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but these cover too small area.
A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation by
approximately 10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5
km).  \cite{zhao2022} reported a similar SW radiation bias in five analysed
CMIP6 models over high-latitude SO and the total cloud fraction underestimation
on the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 (© Jeff Aquilina, Bureau of Meteorology), \textbf{(e)} Vaisala CT25K at
Macquarie Island (© Simon P.\ Alexander, Australian Antarctic Division).
}
\label{fig:map}
\end{figure}

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, the resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her (now His) Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021. Fig. \ref{fig:map} shows a map of the voyages, Table
\ref{tab:voyages} list the voyages, campaigns, and stations, and Table
\ref{tab:voyage-references} lists references where available. Altogether, the
voyages and station dataset comprised 2421 days of data south of 40°S, but the
availability of ceilometer data was slightly shorter due to gaps in
measurements.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
and CT25K, and the Lufft CHM 15k, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar principally designed to  measure cloud base, but they
also measure the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.  We note that
during the MICRE campaign, the ceilometers Vaisala CT25K and CL51 were
installed at the Macquarie Island station concurrently, but in our analysis we
only used the CT25K data obtained from the Atmospheric Radiation Measurement
(ARM) data archive.

@@

The Vaisala CL51 and CT25K (photos in Fig. \ref{fig:map}d, e) are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
The CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K observations. The sampling (temporal) resolution is configurable, and in our
datasets is approximately 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied to the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, there are clear differences particularly in the highest occurence rate regions -- over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results match the observations better than the reanalyses in this metric.

@@

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including peak and total cloud
fraction. In these situations, the model vertical profiles of cloud occurrence compare well with observations, but still tend to
underestimate cloud occurrence above 1 km and near the surface. Non-cyclonic
situations are similar to the general case, partially also because they form
the majority of cases.

When subsetted by relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to relatively unstable situations, where the peak is more diffuse
between 0 and 1 km.  In relatively stable
situations, the models do a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 underestimates cloud occurrence below 1 km (especially near the surface), and
MERRA-2 underestimates cloud occurrence even more strongly.

@@

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in Section \ref{sec:cloud-occurrence}. Observations display the highest proportion of
high cloud cover values (5--8 oktas), peaking at 7 oktas. This pattern is not represented by
ICON or either reanalyses.  While ICON is closest to matching the observed distribution, it tends to be 1 okta
clearer than the observations, peaking at 6 oktas, and substantially underestimating
days with 8 oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

@@

We analysed about 2300 radiosonde profiles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and
temporally colocated profiles were taken from ICON and the reanalyses.
Because the time period covered by the ICON model output (2021--2024) was
different from the time period covered by the observations (2010--2021), when
comparing with the model, we first had to remap the observation time to model
time by taking the same time relative to the start of the year. Consequently,
we also had four virtual/model profiles (one for each year of 2021--2024) for
each observed profile.
The profiles were partitioned into the same subsets as
above (Sections \ref{sec:cloud-occurrence} and \ref{sec:cloud-cover}).  Apart
from relative humidity, we focus on comparing virtual potential temperature
($\theta_v$) due to its role in low-level tropospheric stability, being one of
the primary factors affecting shallow convection and the associated low-level
cloud formation and dissipation. The observed and model profiles of virtual
potential temperature are shown in Fig. \ref{fig:potential-temperature}.

@@

The results imply that SO cloud biases are still a substantial issue even in the
km-scale resolution ICON model, even though an improvement over the
lower-resolution reanalyses is notable. More effort is needed to improve the
model cloud simulations in this understudied region. However, this analysis suggests that the
transition from models with parametrised convection and clouds to
storm-resolving models might not solve these biases without additional effort.
Evaluation of ocean--atmosphere heat, moisture, and momentum fluxes against in-situ observations over the SO and comparison of GSRM simulations against
large-eddy simulations are two potential avenues for future research that could
elucidate the physical mechanisms behind the biases, in addition to the more
common efforts in SO cloud microphysics and precipitation evaluation.
